Approximation: Greedy and Local Search

### 2024-06-22 presentation/presentation.tex
added scheduling on a single machine

\begin{frame}
    \frametitle{Agenda}
    \begin{enumerate}
        \item Introduction
        \item Scheduling Problems
            \begin{itemize}
                \item Scheduling on Single Machines
                \item Scheduling on Multiple Machines 
            \end{itemize}
        \item Graph Problems 
            \begin{itemize}
                \item K-center Problem 
                \item Travelling Salesman Problem
            \end{itemize}
        \item Conclusion
    \end{enumerate}
\end{frame}

@@

\begin{frame}
    \frametitle{Quick Recap: Approximation Ratio}
    $Opt^*$: the optimal solution \\ 
    $Opt$: the suboptimal solution we compute \\[10pt]
    For minimization problem, we have
    \begin{align*}
        \alpha = \dfrac{|Opt|}{|Opt^*|}
    \end{align*}
\end{frame}

\begin{frame}
    \frametitle{Techniques in Approximation}
    \begin{enumerate}
        \item<1-> Randomization: MAXSAT
        \item<2-> Dynamic Programming: Knapsack, Bin Packing
        \item<3-> Linear \& Integer Programming: Primal-Dual, Semidefinite Program 
        \item<4-> \color{red} Greedy \& Local Search \color{black}
    \end{enumerate}
\end{frame}

@@

@@

\begin{frame}{TSP-Limit of Approximation}
    \begin{enumerate}
        \item<1-3> No existent approximation for TSP in general 
        \item<2-3> If there was $\implies$ Hamilton-Cycle solvable in P!
        \item<3> But: metric distance can help us!
    \end{enumerate}
\end{frame}
\pause 
Suppose for each vertice $u, v, w \in V$, triangular inequality holds
\begin{align*}
    d(u, v) + d(v, w) \geq d(u, w)
\end{align*}

@@

\begin{frame}{Algorithm - A bit help from Euler}
    For the size of the tour:
    \begin{enumerate}
        \item $w_{MST} \leq w_{Opt^*}$: Round tour contains trees!
        \item $w_G \leq 2w{MST}$: triangular inequality
    \end{enumerate}
\end{frame}

\begin{frame}{Algorithm - A bit help from Christofide}
\begin{itemize}
    \item We can do the same trick to all Eulerian Graphs
    \item As long as it connects all vertices 
    \item How to make MST Eulerian?
\end{itemize}
\end{frame}

\begin{frame}{Algorithm - A bit help from Christofide}
    Perfect Matching
\end{frame}

>>>>

\begin{frame}
    \frametitle{Agenda}
    \begin{enumerate}
        \item Introduction
        \item Scheduling Problems
            \begin{itemize}
                \item Scheduling with deadlines on a Single machine
                \item Scheduling on Multiple Machines 
            \end{itemize}
        \item Graph Problems 
            \begin{itemize}
                \item K-center Problem 
                \item Travelling Salesman Problem
            \end{itemize}
        \item Conclusion
    \end{enumerate}
\end{frame}

@@

\begin{frame}
    \frametitle{Quick Recap: Approximation Ratio}
    $Opt^*$: the optimal solution \\ 
    $Opt$: the suboptimal solution we compute \\[10pt]
    For minimization problems, we have
    \begin{align*}
        \alpha = \dfrac{|Opt|}{|Opt^*|} > 1
    \end{align*}
    And for maximization problems
    \begin{align*}
        \alpha = \dfrac{|Opt|}{|Opt^*|} < 1
    \end{align*}
\end{frame}

\begin{frame}
    \frametitle{Techniques in Approximation}
    \begin{enumerate}
        \item<1-> Randomization: MAXSAT, MAXCUT
        \item<2-> Dynamic Programming: Knapsack, Bin Packing
        \item<3-> Linear \& Integer Programming: Primal-Dual, Semidefinite Program 
        \item<4-> \color{red} Greedy \& Local Search \color{black}
    \end{enumerate}
\end{frame}

@@

\begin{frame}{Scheduling with deadlines on a Single machine}
    \textbf{Problem Statement: } Given $n$ jobs to be processed on a single machine. 
    How can we schedule them such that they will finish as early as possible. \brk 
    \begin{itemize}
        \item What if there is not deadline?
        \item How do we define earlieness/lateness?
    \end{itemize}
\end{frame}

\begin{frame}{Scheduling with deadlines on a Single machine - Formal Definition}
Suppose job $j$ is finished at time $C_j$, the lateness is 
\begin{align*}
    L_j := C_j - d_j 
\end{align*}
The lateness of all jobs is 
\begin{align*}
    L_{max} = \max_{i \in [n]} L_i
\end{align*}
\textbf{Input:} $n$ jobs with release time $r_j$, processing time $p_j$, and deadline $d_j$. \brk 
\textbf{Output:} a schedule such that $L_{max}$ is minimized.
\end{frame}

\begin{frame}{Algorithm - Intuition}
    If the job $j$ is finished before the deadline we don't get penalty for it, so \dots 
    \begin{enumerate}
        \item<1-> What if we always choose the job with the earliest deadlines 
        \item<2-> But, we may have deal with negative values. 
        \item<3> So all deadlines are negative (by assumption).
    \end{enumerate}
\end{frame}

\begin{frame}{Algorithm - Analysis}
    We start with an observation. 
    Let $S$ be a set of jobs,
    \begin{itemize}
        \item $r(S) := \min_{j \in S} r_j$ 
        \item $p(S) := \sum_{j \in S} p_j$ 
        \item $d(S) := \max_{j \in S} d_j$
    \end{itemize}
    We claim that
    \begin{align*}
        L_{max}^* \geq r(S) + p(S) - d(S)
    \end{align*}
    This is proven by considering the optimal schedule.
\end{frame}

\begin{frame}{Algorithm - Analysis}
    \begin{align*}
        L_{max}^* \geq r(S) + p(S) - d(S)
    \end{align*}
    This leads directly to 
    \begin{align*}
        L_{max}^* \geq r({j}) + p({j}) - d({j}) \geq -d_j
    \end{align*}
    where $j$ is the job that leads to the maximal lateness. 
\end{frame}

\begin{frame}{Algorithm - Analysis}
    \begin{align*}
        L_{max}^* \geq -d_j
    \end{align*}
    Recall $L_{max} = C_j - d_j$, it suffices to show
    \begin{align*}
        L_{max}^* \geq C_j
    \end{align*}
\end{frame}

\begin{frame}{Algorithm - Analysis}
    We start with an ideal scenario: all jobs are released at time $t = 0$. We have 
    \begin{itemize}
        \item $r(S) = 0$ 
        \item $p(S) = C_j$ 
        \item $d(S) < 0$, by assumption
    \end{itemize}
    From lemma it follows that 
    \begin{align*}
        L_{max}^* \geq r(S) + p(S) - d(S) \geq p(S) = C_j
    \end{align*}
    \pause
    \color{red} How about the general cases? \color{black}
\end{frame}

\begin{frame}{Algorithm - Analysis}
    \color{red} How about the general cases? \color{black} 
    \begin{itemize}
        \item<1-> We consider it in a way of calculus
        \item<2-> Find the time $t$ such that $[t, C_j]$ has no idle time 
        \item<3-> Denote jobs processed in this interval with $S$.
        \item<4-> $r(S) = t$, $p(S) = C_j - t$ 
        \item<5> Thus $C_j = r(S) + p(S) \leq L_{max}^*$ 
    \end{itemize}
\end{frame}

\begin{frame}{Algorithm - Analysis} 
    Summarizing two results $L_{max}^* \geq -d_j$ and $L_{max}^* \geq C_j$, we obtain
    \begin{align*}
        L_{max} = C_j - d_j \leq 2L_{max}^*
    \end{align*}
\end{frame}

@@

\begin{frame}{TSP-Limit of Approximation}
    \begin{enumerate}
        \item<1-3> No existent approximation for TSP in general 
        \item<2-3> If there was $\implies$ Hamilton-Cycle solvable in P!
        \item<3> But: metric distance can help us!
    \end{enumerate}
    \pause 
    Suppose for each vertice $u, v, w \in V$, triangular inequality holds
    \begin{align*}
        d(u, v) + d(v, w) \geq d(u, w)
    \end{align*}
\end{frame}

@@

\begin{frame}{Algorithm - A bit help from Euler}
    For the size of the tour:
    \begin{enumerate}
        \item $w_{MST} \leq w_{Opt^*}$: Round tour contains trees!
        \item $w_G \leq 2w{MST}$: triangular inequality
    \end{enumerate}
    $\implies w_G \leq 2w_{Opt^*}$ 
\end{frame}

\begin{frame}{Algorithm - A bit help from Christofide}
\begin{itemize}
    \item<1-> We can do the same trick to all Eulerian Graphs
    \item<2-> As long as it connects all vertices 
    \item<3-> How to make MST Eulerian?
    \item<4-> Make all vertices have \textbf{even} degrees
\end{itemize}
\end{frame}

\begin{frame}{Algorithm - A bit help from Christofide}
    \textbf{HOW?}
    \begin{enumerate}
        \item<1-> Connect all vertices with odd degrees
        \item<2-> How many vertices with odd degrees?
        \item<3-> Finding a perfect matching
        \item<4-> What is the cost? What is the size of the PM?
    \end{enumerate}
\end{frame}

\begin{frame}{Algorithm - A bit help from Christofide}
    For the size of the tour:
    \begin{enumerate}
        \item $w_{MST} \leq w_{Opt^*}$: see 1st part
        \item $w_{PM} \leq \dfrac{1}{2} w_{Opt^*}$:
        \item $w_{G} \leq w_{MST} + w{PM}$: traingular inequality
    \end{enumerate}
    $\implies w_{G} \leq \dfrac{3}{2} w_{Opt^*}$
\end{frame}
